---
title: "FIESTA - Analysis Functions - Core Tables/Reports"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FIESTA - Cores Tables/Reports}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r setup, include = F}
library(knitr)
knitr::opts_chunk$set(message = F, warning = F)
```

```{r, include=FALSE}
# Sets up output folding
hooks = knitr::knit_hooks$get()
hook_foldable = function(type) {
  force(type)
  function(x, options) {
    res = hooks[[type]](x, options)
    
    if (isFALSE(options[[paste0("fold.", type)]])) return(res)
    
    paste0(
      "<details><summary>", type, "</summary>\n\n",
      res,
      "\n\n</details>"
    )
  }
}
knitr::knit_hooks$set(
  output = hook_foldable("output"),
  plot = hook_foldable("plot")
)
```


## Core Tables/Reports Overview

Estimates generated from `FIESTA` are typically reported for a single FIA Evaluation. But we can use `FIESTA's` Analysis functions to generate a set of tables each time or use `FIESTA's` RMarkdown templates to generate a standard report each time.

The main objective of this tutorial is to demonstrate the use of `FIESTA` Analysis functions for generating a core set of tables and a core report for a given list object of population data, resulting from the `modGBpop` function. 


### Example Data

The examples use FIA plot data from all Evaluations in Vermont using the FIA annual plot design, downloaded and stored in a SQLite database on May 13, 2022, as external data in `FIESTA`. The following code was used to create the database.

## Get data from datamart and add to SQLite database
SQLitefn <- "F:/FIAdata/FIApublic_VT_eval1519_pop_20221213.db"
DBgetPlots(istree=TRUE, isseed=TRUE, 
          states = "Vermont", 
          invtype = "ANNUAL",
          eval = "FIA", 
          eval_opts = list(Endyr = 2015:2019, evalType = "VOL"),
          getxy = FALSE,
          returndata = FALSE, 
          savePOP = TRUE, 
          othertables = c("POP_STRATUM", "POP_ESTN_UNIT"),
          savedata = TRUE, 
          savedata_opts = list(out_fmt = "sqlite", 
                               out_dsn = SQLitefn, 
                               overwrite_dsn = TRUE)
          )

```{r}
sqlitefn <- system.file("extdata", "FIA_data/FIApublic_VT_eval1519_pop_20221213.db", package="FIESTAnalysis")

conn <- DBI::dbConnect(RSQLite::SQLite(), sqlitefn)
DBI::dbListTables(conn)
pop_stratum <- DBI::dbReadTable(conn, "pop_stratum")
sort(unique(pop_stratum$EVALID))

DBI::dbDisconnect(conn)

```


### Set up

First, you'll need to load the necessary libraries:

```{r, warning = F, message = F}
library(FIESTA)
library(FIESTAnalysis)
library(data.table)
```


Next, you'll need to set up an "outfolder". This is just a file path to a folder where you'd like `FIESTA` to send your data output. For this vignette, we have saved our outfolder file path as the `outfolder` object in a temporary directory. We also set a few default options preferred for this vignette.

```{r}
outfolder <- tempdir()
#outfolder <- "E:/workspace/FIESTA/FIESTA_core"
```




Now, generate the population data for the state of Vermont using the example database described above. 
```{r, GBpopdat}
 
GBpopdat <- modGBpop(popTabs = list(plt="plot", cond="cond", tree="tree", seed="seed"),
                     dsn = sqlitefn,
                     pltassgn = "pop_plot_stratum_assgn",
                     pltassgnid = "PLT_CN", 
                     popFilter = list(evalid = 501901),
                     areavar = "AREA_USED",
                     unitarea = "pop_estn_unit",
                     unitvar = "ESTN_UNIT",
                     stratalut = "pop_stratum",
                     strata_opts = list(getwt = TRUE, 
                                      getwtvar = "P1POINTCNT")
                     )
                                
```


Next, using the population data above, generate a set of core tables to save to outfolder.

```{r}

tabfolder <- file.path(outfolder, "GBtables")
if (!dir.exists(tabfolder)) {
  dir.create(tabfolder)
}
anGBpop_core(GBpopdat = GBpopdat, 
             title.ref = "Vermont, eval2019", 
		         outfolder =tabfolder
             )

```


Next, using the population data above, generate a core report and save to outfolder.

```{r}

anGBpop_report(GBpopdat = GBpopdat, 
               AOInm = "Vermont",
               title.ref = "Vermont, eval2019", 
		           outfolder = outfolder
               )

```


Now, let's generate the population data for a county in Wyoming using the data stored in FIESTA. 
```{r, MApopdat}

unit <- 1

MApopdat <- modMApop(popTabs = list(plt=WYplt, cond=WYcond, tree=WYtree, seed=WYseed),
                   pltassgn = WYpltassgn[WYpltassgn$ESTN_UNIT == unit, ],
                   pltassgnid = "CN", 
                   areavar = "ACRES",
                   unitarea = WYunitarea[WYunitarea$ESTN_UNIT == unit, ],
                   unitvar = "ESTN_UNIT",
                   unitzonal = WYunitzonal[WYunitarea$ESTN_UNIT == unit, ],
                   prednames = c("dem", "tcc", "tpi", "tnt"),
                   predfac = "tnt"
                   )
                                
```

Next, using the population data above, generate a set of core tables to save to outfolder.

```{r}

tabfolder <- file.path(outfolder, paste0("MAtables_", unit))
if (!dir.exists(tabfolder)) {
  dir.create(tabfolder)
}
anMApop_core(MApopdat = MApopdat, 
             MAmethod = "greg",
             title.ref = "Wyoming, eval2013", 
		         outfolder =tabfolder
             )

```


Next, using the population data above, generate a core report and save to outfolder.

```{r}

anMApop_report(MApopdat = MApopdat, 
               AOInm = "Wyoming-County1",
               title.ref = "Wyoming, eval2013", 
		           outfolder = tabfolder
               )

```